---
title: 'W7: Homework'
author: "Helene Andersen"
date: "20 July 2020, updated `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal
I want to be able to analyse the data from OECD on gender inequality. My focus will be on the share of female seats at the largest listed companies. Data from: https://stats.oecd.org/index.aspx?queryid=54753 

# Solution
First, install a handful of classic R packages and load their libraries:

- `rvest` for web-scraping
- `dplyr` for data-wrangling
- `tidyr` for data transformation
- `stringr` for string manipulation
- `janitor` for clean headers that your OCD will love you for


```{r libraries}
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
```


## Scrape the data

Next, learn how scrape the content of the website and extract the HTML table:
```{r}
url <- "https://stats.oecd.org/index.aspx?queryid=54753"
# scrape the website
url_html <- read_html(url)
```

The data have to be cleaned before I can use it

```{r , option1, eval=FALSE}
whole_table <- url_html %>% 
 html_nodes("tr") %>% #en række (tr)
 html_text(trim = FALSE) 
head(whole_table)
```
I have to sort out the first lines of the table for it to make sense

```{r, option2}
whole_table <- url_html %>% 
 html_nodes("td") %>% #en individuel celle (td)
 html_text(trim = FALSE) 
head(whole_table)
```


```{r, option3}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list

```

If you run a `head()` on the resulting `whole_table`, you will see that it is a list with unnamed elements marked by numbers in double brackets, such as [[1]]. 
What is a list? A table-like looking structure that needs one more transformation (in this case, unlisting) before it becomes a useful dataframe. 

```{r htmltodf}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table)
```

Now I tidy the data by saying it should only count in the rows 8-51 nand the colomns 2-13

```{r}
femaleseats <- new_table[8:51,2:13] #chose the rows and columns I want to look at
femaleseats <- femaleseats[,c(1,3:12)] #chose the columns I want to eliminate/look at
colnames(femaleseats) <- c("Country", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019")
```

Now I will make some visualization on my script

```{r}
fseats <- as_tibble(femaleseats) 
```

```{r}
fseats %>% 
	filter(Country== "Denmark")
```

```{r}
plot(fseats, "Denmark")
```
This did not work as the visualization.... but I did not want to spend any more time on it!

