---
title: "R for Anthropologists"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#INTRODUCTION
In this guide we want to investigate the question: "How does the length of maternity leave influence the share of female managers in the Nordic Region?". Our process of this investigation will work as a guide on how to compare and see correlations between to quantitative datasets from OECD.Stat.Following this guide you could choose investigate and visualize other data sets. Our guide includes 3 steps:

1. FIRST DATASET: Scrape, tidy, code and visualize statistical data from the dataset "Employment: Length of maternity leave, parental leave, and paid father-specific leave"

2. SECOND DATASET: Scrape, tidy, code and visualize statistical data from the dataset "Employment: Share of female managers"

3. MERGE DATASETS: Merge the two datasets to see if there is a correlation between the length of maternity leave and the share of female managers in the Nordic Region

# LOADING PACKAGES
To get started we need to load different packages that will be useful in our script and examination. Each package allows us to access the data in different ways.

```{r libraries}
library(rvest) #to scrape information from web pages
library(dplyr) #to abstract over how the data is stored
library(tidyr) #to create tidy data
library(stringr) #to work with strings as easy as possible
library(janitor) #to examine and clean dirty data
library(tidyverse) #to collect some of the most versatile R packages
library(ggplot2) #to create visualizations and graphs
```

# 1: SCRAPING FIRST DATA FROM OECD.STAT
To investigate the datasets from OECD.Stat we will now scrape the website for data. I scrape the content of the website and extract the HTML table
```{r webscrape}
url <- "https://stats.oecd.org/index.aspx?queryid=54760&fbclid=IwAR1Jt-NDjC_e4iBKlxrIpKSIBGtmsJYY1V2R6UmEIXT_sAelAk1_2S_YvRU#" 
# scrape the website. Here you insert the link of the data set of interest
url_html <- read_html(url)
```

We extract the whole HTML table through the <table> tag.
```{r webscrape and assign object}
whole_table <- url_html %>% #we assign the data to a new object
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list
```

If you run a `head()` on the resulting `whole_table`, you will see that it is a list with unnamed elements marked by numbers in double brackets, such as [[1]]. 
```{r}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table) 
```

The line above takes the whole table from the website with all rows and column. This is very difficult to read. It does not look nice and it shows things that we are not interested in. Therefore we tidy the data in the following section of this guide.

# Tidying the data

In our case we only want to look at the countries' average of weeks of maternity leave through the time period (1990 to 2016). Therefore we filter the data as shown in the following lines

```{r, create the table}
new_table <- as.data.frame(new_table) #create a new table
mleave <- new_table[8:44,2:30] #choose the rows and columns we want to look at 
mleave <- mleave[,c(1,3:29)] #choose the columns we want to both eliminate and look at
colnames(mleave) <- c("country", "1990", "1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016") #assign new column names to years

mleave <- as.data.frame(lapply(mleave, function(y) gsub("\\.\\.", NA, y))) #change the .. to NA, så R can read the data

mleave #showing the table
```

In the table the numbers are set at characters. In the following I change these to numeric, so that R can make read the table correctly.

```{r, numbers to numeric}
cols.num <- c("X1990", "X1991", "X1992", "X1993", "X1994", "X1995", "X1996", "X1997", "X1998", "X1999", "X2000", "X2001", "X2002", "X2003", "X2004", "X2005", "X2006", "X2007", "X2008", "X2009", "X2010", "X2011", "X2012", "X2013", "X2014", "X2015", "X2016") #columns from characters to numeric
mleave[cols.num] <- sapply(mleave[cols.num],as.numeric) #recode characters as numeric
sapply(mleave, class) #print classes of all colums

```

In the table the countries are set to be characters. In the following I change these to factors.

```{r, countries from character to factor}
mleave$country <- as.factor(mleave$country) #columns from characters to factor
```

In the following I restructure the table so it knows that every country has numbers from different years.

```{r, years}
mleave_long <- mleave %>% gather(year, weeks_mleave, -country) #restructuring the table so it knows that every country has numbers from different years

```

Now the table is readable for R.

# VISUALIZATIONS
Now that we have finished tidying the table from OECD.Stat, we can begin making visualizations of this independent data set. In order to do so we need a few more packages.

```{r, packages}
install.packages("gifski")
install.packages("png")
library(gifski) #to convert images to GIF animations
library(png) #to read, write and display the PNG format
library(scales) #to manipulate and polish visualizations
library(theme)
library("ggthemes") #to make nice visualizations
library(gganimate) #to make animated visualizations
```

There are several ways of visualizing data. You could e.g. look up line geoms, point geoms, bars geoms. One way of vizualizing data is the following. 

```{r, geom jitter}
mleave_long %>% 
  ggplot(aes(x = year, y = weeks_mleave))+
  geom_jitter(aes(color = country))
```
Another way of visualizing the data can be done like this:

```{r, geom line}
ggplot(mleave_long) +
  geom_line(mapping = aes(x = year, y = weeks_mleave, group = country, color = country))
```
As we can see above, many years are included in the visualization. We want to make our visualizations more clear and specific by focusing on the countries in the Nordic Region: Denmark, Iceland, Sweden, Norway and Finland. We chose to compare these countries, as they have similiar scandinavian welfare models. Furthermore, since the dataset which we want to compare this with only have data from 2011-2018, we filter the years so our new assigned value will include the countries 2011-2016. 

```{r assigning the nordic region}
mnordic <-  mleave_long %>% filter(country %in% c("Denmark","Iceland", "Sweden","Norway", "Finland")) %>%  filter(year %in% c("X2011", "X2012", "X2013", "X2013", "X2014", "X2015", "X2016")) #assigning the nordic countries and time period
mnordic$year <- str_remove(mnordic$year, "X") #filtering out the "X" that popped in to our model earlier
```

Now that we have created this object with the specific countries we will use it to visualize the differences within the countries. We will put it in the earlier ggplot-visualization and add a title.
```{r}
ggplot(mnordic) +
  geom_line(mapping = aes(x = year, y = weeks_mleave, group = country, color = country)) + ggtitle("Length of Maternity Leave in Nordic Countries") +theme_solarized()+ scale_color_solarized()+ #add a title
  xlab("Year")+ #naming the x axis
  ylab("Weeks") #naming the y axis
```

Here we used the theme solarized. You could choose among many different from this site: https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/

The visualization above shows the length of maternity leave. With this script it is possible to compare the different countries and to select specific countries you want to compare.


# 2: SCRAPING SECOND DATA FROM OECD.STAT

# Scraping Data 

You can now choose another dataset that you would like to work with. Here we scrape the content of the website and extract the HTML table:
```{r}
url <- "https://stats.oecd.org/index.aspx?queryid=96330"
# scrape the website
url_html <- read_html(url)
```

The following option let's us extract the whole HTML table through the <table> tag.
```{r, option3}
whole_table <- url_html %>% 
 html_nodes("table") %>%
 html_table(fill = TRUE)  #str(whole_table) turns out to be a list
```

If you run a `head()` on the resulting `whole_table`, you will see that it is a list with unnamed elements marked by numbers in double brackets, such as [[1]]. 
```{r htmltodf}
new_table <- do.call(cbind,unlist(whole_table, recursive = FALSE)) 
head(new_table) 
```

The line above takes the whole table from the website with some rows and coloumns. It does not look nice and it shows thing that I am not interested in. 

# Tidying the data
In our case we only want to look at the countries' share of female managers through the time period (2011 to 2018). Therefore we filter the data as shown in the following lines

```{r create the table}
fmanagers <- new_table[8:52,2:18] #choose the rows and columns we want to look at 
fmanagers <- fmanagers[,c(1,3:10)] #choose the columns we want to both eliminate and look at
colnames(fmanagers) <- c("country", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018") #assign new column names to years
```

I can see that the data is listed in characters (<chr>), and I want to convert some of it to numeric to be able to create visualizations (I got help from this website: https://stackoverflow.com/questions/22772279/converting-multiple-columns-from-character-to-numeric-format-in-r)

```{r from characters to numeric}
cols.num <- c("2011","2012","2013","2014","2015","2016","2017","2018") #we tell r which colons should be numeric
fmanagers[cols.num] <- sapply(fmanagers[cols.num],as.numeric) #in this section we convert the colomns in our dataset to be numeric instead of characters
sapply(fmanagers, class) #now we use the 'sapply' to check the colomns' classes. We see that now it is numeric, and therefore we can begin to visualize our data
```

To be able to create the wanted visualizations, I will need to assign the different countries as factors so that I can distinguish between them. In the following I will ascribe the countries as factors:

```{r countries from character to factor}
fmanagers$country <- as.character(fmanagers$country)
sapply(fmanagers, class)
```

In the following we restructure the table so that R can read it.

```{r, the countries with highest shares in 2011}
fmanagers_long <- fmanagers %>% gather(year, percentage, -country) # I want to tidy my data further so that there will be a specific column with the year assigned. Now I can hopefully start visualizing my data...
```

Now we have finished tidying the data and restructuring the table. Therefore we will now be able to move on with the visualizations

# Visualizations of second data set

```{r}
theme_set(theme_bw()) #set the theme to black and white for better visualization
ggplot(fmanagers_long) +
  geom_line(mapping = aes(x = year, y =percentage, group = country)) #by selecting the country and mapping that the visualization should include years, countries and percentage
```
Here we have all the years and all the countries listed. But it does not look very manageable, so we will try to pick out the countries we want to compare and illustrate them in different colors. The chosen countries are: Denmark, Iceland, Colombia, Italy, and the OECD average.

```{r}
fmanagers_long %>% 
  ggplot(aes(x =  year, y = percentage))+
  geom_jitter(aes(color = country))
```

Another way of making visualizations is the following:

```{r}
ggplot(fmanagers_long) +
  geom_line(mapping = aes(x =  year, y = percentage, group = country, color = country))
```
As there are so many countries in the data set, the vizualization is not very nice. Therefore we now want to select some countries. We choose five countries; Denmark, Iceland, Finland, Norway and Sweden to see how the share of female managers and the development in these countries look. Therefore we create a new object called 'fnordic'.

```{r}
fnordic <-  fmanagers_long %>% filter(country %in% c("Denmark","Iceland", "Sweden","Norway", "Finland")) %>% filter(year %in% c("2011","2012", "2013","2014", "2015", "2016")) #here you can chose the countries and the years you want to focus on
```

Now that we have created this object with the specific countries we will use it to visualize the differences within the countries. We will put it in the earlier ggplot-visualization and add a title.

```{r}
ggplot(fnordic) +
  geom_line(mapping = aes(x =  year, y = percentage, group = country, color = country)) + ggtitle("Share of Female Managers in Nordic Countries")+ theme_solarized()+ scale_color_solarized()+ #add title
  xlab("Year")+ #to rename the x-axis
  ylab("Percentage")#to rename the y-axis
```

Here we used the theme solarized. You could choose among many different from this site: https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/

The visualization above shows the share of female managers. With this script it is possible to compare the different countries and to select specific countries you want to compare. As we can see above, the countries that Denmark normally compare it self with are doing a lot better than Denmark when it comes to the share of female managers. Also, Denmark (and Finland) is doing worse than the OECD average which includes the 45 OECD countries.

To make it easier to follow, we will repost the visualization showing the length of maternity leave in the Nordic Region. 

```{r}
ggplot(mnordic) +
  geom_line(mapping = aes(x = year, y = weeks_mleave, group = country, color = country)) + ggtitle("Length of Maternity Leave in Nordic Countries") +theme_solarized()+ scale_color_solarized()+ #add a title
  xlab("Year")+ #naming the x axis
  ylab("Weeks") #naming the y axis
```
Now, that we can see these two visualizations showing the development from 2011-2016 in the Nordic Region, we want to show how it is possible to merge the two datasets/visualizations!


# MERGING THE DATASETS AND CREATING ANIMATION

We have worked with two data sets from OECD.Stat independently. In the following section we merge the two data sets to and create an animation. We ask the question: How does the length of maternity leave influence the share of female managers in the Nordic countries? 

Firstly we put the data sets together. We do this by the function 'merged'. You want to make sure that your two datasets have the same variables and rows to compare them.
```{r merged dataset}
merged <- merge(mnordic, fnordic, by=c("year", "country"))
```

Then we will try to make visualizations on this. At first, we make the comparisons within a certain year, for instance 2016.

```{r, merged with years}
# I will try to merge the two statistics with specifik years

merged2016 <- filter(merged, year == "2016") #we can compare the different countries on maternity leave and share of female managers in 2016

# Here, we will visualize the status in 2016 on the length of maternity leave and share of female managers in the Nordic Region. We call this animation 'anim1', but you could also call it something else

anim1 <- ggplot(merged2016,aes(x = weeks_mleave, y = percentage, group = country, color = country))+
  geom_point(size=8, alpha = 0.5)+
  xlim(10,24)+
  ylim(25,41)+
  xlab("Weeks")+
  ylab("Percentage")+
  ggtitle("MATERNITY LEAVES' INFLUENCE ON SHARE OF FEMALE MANAGERS")+
  geom_text(aes(label=country),hjust=0, vjust=2.5)+theme_solarized()+ scale_color_solarized()+
  theme(legend.position = "none")

anim1 
```

Now we want to make a moving visualization with changing years.

```{r, animation}
class(merged$year) #the class of year is character, so we need to change it
merged$year <- as.numeric(as.character(merged$year)) #we change 'year' to numeric, so the visualization will be able to move through the years as numbers

anim2 <-ggplot(merged, aes(x = weeks_mleave, y = percentage, group = country, color = country))  +
  geom_point(size=8, alpha = 0.4)+
  transition_time(year)+
   xlab("Weeks of maternity leave")+
  ylab("Percentage of female managers")+
  labs(title = "Year: {round(frame_time)}")+
  theme_solarized()
anim2 #we call this visualization anim2, but it could also have a different name
```

Now we have maked the data move! As we can see, Denmark is the country with the longest length of maternity leave and the lowest share of female managers. However it does not look at if there is a direct correlation between the length of maternity leave and the share of female managers.   


If you want to use this animation in powerpoints or other kinds of presentations, it is a good idea to save it as a GIF. Below is a guide on that:

```{r}
#Make it a GIF
animate(anim2, fps = 10, width = 750, height = 450) #we want to make it a different format
anim_save("gender.gif")
```

# WRAP IT UP    
Now, we have learned how to scrape, code, tidy and visualize data from OECD.Stat. Good job! We hop you can use this in your future research. If you want to contact us, our GitHubs are: "helenefonne" and "sofia-sif".



